---
title: "Peer Assessments - Project 1"
output: html_document
---

This is the Peer Assessment from project 1 of course Reprodutive Research.

The data for this assignment is "Activity monitoring data". The variables included in this dataset are:

**steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)

**date**: The date on which the measurement was taken in YYYY-MM-DD format

**interval**: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

The following activities are archeive:

##A)Loading and preprocessing the data


```{r, echo=TRUE}
setwd("~/projeto MINERACAO DE DADOS/Aula Coursera Data Mining/5 - Reproducible Researsh/Peer Assessments/Peer Assessments 1")
x<-read.csv("activity.csv", header = TRUE, sep = ",", quote = "\"",
            dec = ".", fill = TRUE, comment.char = "",stringsAsFactors = FALSE)
```

##B)What is mean total number of steps taken per day?

###Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}

library(dplyr)
grupo_data<-group_by(x,date)
tot_day<-summarize(grupo_data,total=sum(steps,na.rm=TRUE))
qtd<-tot_day$total
hist(qtd, col="red", main="Total number of steps", xlab="Steps")

dev.copy(png,'histogram1.png', width=480, height=480, units="px")
dev.off()
```


###Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}

mean<-mean(tot_day$total,na.rm=TRUE)
median<-median(tot_day$total,na.rm=TRUE)

```

Mean: `r mean` 

Median: `r median` 


##C)What is the average daily activity pattern?

###Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days


```{r, echo=TRUE}

grupo_data<-group_by(x,interval)
tot_interv<-summarize(grupo_data,mean=mean(steps,na.rm=TRUE))

plot(tot_interv$interval,tot_interv$mean,type="l", col="blue",main="Mean of steps", ylab="Steps",xlab="inteval")

dev.copy(png,'time_series_plot.png', width=480, height=480, units="px")
dev.off()

```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE}
max_steps<-max(tot_interv$mean)
five_minutes<-tot_interv[tot_interv$mean==max_steps,1]

```

The maximun number of steps is `r max_steps` and occurs in the `r five_minutes` interval.



##D) Imputing missing values

###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo=TRUE}
missing<-x[is.na(x$steps),]
n_missing<-nrow(missing)

```

The total number of missing values is `r n_missing`.


###Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r, echo=TRUE}
complete_data<-merge(x,tot_interv)
complete_data$step2<-ifelse(is.na(complete_data$step),complete_data$mean,complete_data$step)
new_data_set<-complete_data[,c("interval","date","step2")]
names(new_data_set)[3]<-"steps"
new_data_set$datePOSIX<-strptime(new_data_set$date,"%Y-%m-%d")

```

###Make a histogram of the total number of steps taken each day 


```{r, echo=TRUE}

grupo_data2<-group_by(new_data_set[,c("date","steps")],date)
tot_day2<-summarize(grupo_data2,total=sum(steps,na.rm=TRUE))
qtd2<-tot_day2$total

par(mfrow=c(1,2))
hist(qtd2, col="red", main="Total number of steps without NA", xlab="Steps")
hist(qtd, col="red", main="Total number of steps with NA", xlab="Steps")

dev.copy(png,'histogram2', width=480, height=480, units="px")
dev.off()

```



###Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}

mean2<-mean(tot_day2$total,na.rm=TRUE)
median2<-median(tot_day2$total,na.rm=TRUE)

```


Mean with NA: `r mean` 

Median with NA: `r median` 


Mean without NA: `r mean2` 

Median without NA: `r median2` 


Impact on Mean: `r mean2-mean`

Impact on Median: `r median2-median` 

##Are there differences in activity patterns between weekdays and weekends?

###Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}

new_data_set$week<-factor(new_data_set$datePOSIX$wday %in% c(0,6),labels=c("weekday","weekend"))

```

###Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using simulated data:


```{r, echo=TRUE}
#Select Weekday
WeekdayData<-new_data_set[new_data_set$week=="weekday",]
WeekendData<-new_data_set[new_data_set$week=="weekend",]

grupo<-group_by(WeekdayData[,c("interval","steps")],interval)
WeekdaySteps<-summarize(grupo,mean=mean(steps,na.rm=TRUE))

grupo<-group_by(WeekendData[,c("interval","steps")],interval)
WeekendSteps<-summarize(grupo,mean=mean(steps,na.rm=TRUE))

par(mfrow=c(2,1))

plot(WeekdaySteps$interval,WeekdaySteps$mean,type="l", col="blue",main="Mean of steps in Weekdays", ylab="Steps",xlab="inteval")

plot(WeekendSteps$interval,WeekendSteps$mean,type="l", col="blue",main="Mean of steps in Weekends", ylab="Steps",xlab="inteval")


dev.copy(png,'panel_plot.png', width=480, height=480, units="px")
dev.off()
```
